\texttt{clarify}: Simulation-Based Inference for Regression Models

### 2025-09-18 rjournal_paper/Submission 2/clarify.tex
File management

\abstract{%
Simulation-based inference is an alternative to the delta method for computing the uncertainty around regression post-estimation (i.e., derived) quantities such as average marginal effects, average adjusted predictions, and other functions of model parameters. It works by drawing model parameters from their joint distribution and estimating quantities of interest from each set of simulated values, which form a simulated ``posterior'' distribution of the quantity from which confidence intervals can be computed. clarify provides a simple, unified interface for performing simulation-based inference for any user-specified derived quantities as well as wrappers for common quantities of interest. clarify supports a large and growing number of models through its interface with the marginaleffects package and provides native support for multiply imputed data.
}

@@

The delta method involves computing a first-order Taylor series approximation to the variance of the derived quantity, and standard Wald-based inference relies on computing quantiles based on the Normal distribution and using them to compute p-values and confidence intervals. \CRANpkg{clarify} implements an alternative to the delta method---simulation-based inference---which involves simulating a ``posterior'' distribution of the derived quantities. Simulation-based inference does not require understanding Taylor series or the calculus that underlies it, which can make it more palatable to nontechnical audiences and easier to learn for students without necessarily sacrificing statistical performance \citep{kingMakingMostStatistical2000, zelnerUsingSimulationInterpret2009}. Some studies have found that simulation-based inference performs as well or better than the delta method for computing derived quantities (i.e., with respect to achieving close to nominal coverage for confidence intervals), especially for complicated derived quantities and in smaller samples \citep{mackinnonConfidenceLimitsIndirect2004, holeComparisonApproachesEstimating2007, herronPostestimationUncertaintyLimited1999}. Its empirical performance has been particularly well studied in the context of mediation analysis, in which the quantities of interest are products and ratios of regression coefficients, where it has been shown to perform well relative to the delta method due to the non-Normality of these quantities \citep{tofighiMonteCarloConfidence2016, preacherAdvantagesMonteCarlo2012}.

The methodology \CRANpkg{clarify} relies on was developed by \citet{krinsky1986} and is described in \citet{kingMakingMostStatistical2000} and \citet{herronPostestimationUncertaintyLimited1999}. Simulation-based inference involves taking draws from a specified joint distribution of model parameters, computing derived quantities from these draws, and collecting the derived quantities in a ``posterior'' distribution, from which uncertainty measures (standard errors and confidence intervals) can be computed. This method assumes the model parameters are drawn from a multivariate Normal (or T) distribution with means at the estimated values and covariance equal to the asymptotic covariance matrix of the estimated values, a standard assumption motivated by the central limit theorem that underlies usual inference on the original model parameters. Arriving at the posterior distribution does not require taking any derivatives or making any approximations beyond those usually used for inference on model parameter estimates, except for the approximation due to Monte Carlo error induced by sampling from a finite number of simulations (which can always be reduced by increasing the number of draws at the cost of increased computing time).

The nonparametric bootstrap is another alternative to the delta method for inference that does not require its analytic approximations \citep{efronBootstrapMethodsStandard1986}; bootstrapping typically involves re-sampling individuals from the sample, fitting the model in each bootstrap sample, and computing the quantity of interest from each model. Although bootstrapping tends to work well in practice, especially for complex and non-Normal estimators, refitting the model repeatedly can be prohibitively time-consuming and computationally expensive, especially for complicated models or large datasets. Simulation-based inference only requires the model to be fit once, and the simulations involve taking draws from a distribution produced from the single set of estimated parameters, making it much quicker in practice and allowing the user to capitalize on the already valid estimation of the model parameters. Methods for computing valid confidence intervals in cases when the quantity of interest has a complicated distribution are better developed when bootstrapping, however \citep{efronBootstrapMethodsStandard1986}.

More formally, we fit a regression model \(y_i = f(x_i; \beta)\), such as a linear or other generalized linear model with model coefficients \(\beta\). We assume \(\hat{\beta} \sim \text{MVN}(\beta, \Sigma_{\hat{\beta}})\) where \(\hat{\beta}\) is the vector of estimates of \(\beta\) and \(\Sigma_{\hat{\beta}}\) is their asymptotic covariance matrix. We define a function \(\tau(\beta)\) that represents a quantity of interest derived from the model parameters, and compute its estimate \(\widehat{\tau(\beta)}\) as \(\tau(\hat{\beta})\).

To perform simulation-based inference, we take \(M\) draws \(\tilde{\beta}^{(j)}\) for \(j\in(1, \dots, M)\) from a multivariate Normal distribution with mean vector \(\mu = \hat{\beta}\) and covariance \(\Sigma = \hat{\Sigma}_{\hat{\beta}}\), where \(\hat{\Sigma}_{\hat{\beta}}\) is an estimate of the asymptotic covariance matrix of the parameter estimates. We use the distribution of \(\tilde{\tau}=\tau(\tilde{\beta})\) as the ``posterior'' distribution of \(\widehat{\tau(\beta)}\), and compute its variance as
\begin{equation}
\hat\sigma^2_{\widehat{\tau(\beta)}} = \frac{1}{M-1}\sum^M_{j=1}{(\tilde{\tau}^{(j)} - \bar{\tilde{\tau}})^2}
\label{eq:posteriorvariance}
\end{equation}
and quantile \(100(1-\alpha)\%\) confidence interval limits as \(\left[\tilde{\tau}_{(\frac{\alpha}{2})}, \tilde{\tau}_{(1-\frac{\alpha}{2})}\right]\) where \(\tilde{\tau}_{(q)}\) is the \(q\)th value of \(\tilde{\tau}\) when arranged in ascending order (i.e., the \(q\)th quantile of the empirical cumulative distribution function of \(\tilde{\tau}\)). Simulation-based Wald-type confidence intervals can be computed as
\begin{equation}
\left[\tau(\hat{\beta}) + \hat\sigma_{\widehat{\tau(\beta)}} Z_{\frac{\alpha}{2}}, \tau(\hat{\beta}) + \hat\sigma_{\widehat{\tau(\beta)}} Z_{1-\frac{\alpha}{2}} \right]
\label{eq:siminterval}
\end{equation}
where \(Z_q\) is the \(q\)th quantile of a standard Normal distribution. The delta method-based Wald-type confidence intervals use this formula but with the first-order Taylor approximation to the asymptotic variance:
\begin{equation}
\hat\sigma^2_{\widehat{\tau(\beta)}}=\nabla\tau(\hat{\beta}) \Sigma_{\hat{\beta}} \nabla\tau'(\hat{\beta})
\label{eq:deltavariance}
\end{equation}
where \(\nabla\tau(\hat{\beta})\) is the gradient of \(\tau(\beta)\) with respect to \(\beta\) evaluated at \(\hat{\beta}\).

To compute a p-value for a hypothesis test involving the quantity of interest, i.e., \(\text{H}_0: \tau (\beta) = \tau_0\) with a given null value \(\tau_0\), we can invert the confidence interval \citep{thulinModernStatisticsWrangling2021}; that is, we find the largest value of \(\alpha\) such that \(\tau_0\) is within the confidence interval and use that \(\alpha\) as the p-value for the test. For Wald-based inference (either using the simulation-based variance or delta method-based variance), this is equivalent to performing a standard two-sided Z-test using the test statistic
\begin{equation}
Z=\frac{\tau(\hat\beta)-\tau_0}{\hat\sigma_{\widehat{\tau(\beta)}}}
\label{eq:waldz}
\end{equation}
One benefit of using the quantile p-values for inference is that equivalent tests of the same hypothesis will always yield identical p-values; for example, testing the equality of two derived quantities will yield the same p-value when comparing the difference between the quantities against a null hypothesis of 0 and the ratio of the quantities against a null hypothesis of 1, as each of these hypothesis is true if and only if the other is true.

One would expect simulation-based quantile inference, simulation-based Wald inference, and delta method-based Wald inference to align when the posterior is Normally distributed around the estimate, in which case any discrepancies would be due to Monte Carlo error in the simulated values (and therefore would shrink with increasing draws). However, for low values of \(\alpha\), it may require many draws for the simulation-based intervals to stabilize; delta method-based intervals are not subject to this error. There are a few cases in which the results might diverge: in some cases, the first-order Taylor series approximation to the variance may be poor, though in practice the approximation error is small and shrinks quickly with increasing sample size. When the posterior distribution is non-Normal but symmetric around the estimate, the quantile intervals may be more accurate (i.e., in the sense of achieving closer to nominal coverage) because they do not rely on quantiles from the Normal distribution \citep{tofighiMonteCarloConfidence2016}.

Another potential advantage quantile intervals can have over Wald-intervals is that when \emph{some} monotonic transformation of the estimate has a symmetric distribution centered around the transformed estimate, the quantile intervals can achieve correct coverage without requiring knowledge of \emph{which} transformation is required \citep{efronBootstrapMethodsStandard1986}; this is true of the quantile-based p-values as well. When the distribution is not centered around the estimate and no monotonic transformation will make it so, though, neither quantile-based nor Wald-based intervals would be expected to perform well, and quantile intervals could yield even worse coverage than Wald-based intervals, a phenomenon that occurs in the context of bootstrapping \citep{efronBootstrapMethodsStandard1986}\footnote{We thank an anonymous reviewer for pointing out a scenario in which this could occur: for a quantity of interest with a right-skewed sampling distribution, one would prefer an estimate to the right of the quantity's true value to have a confidence interval skewed to left to capture the bulk of the sampling distribution, but in practice a quantile confidence interval would also be skewed to the right. While a symmetric Wald-based interval may not have adequate coverage, the quantile-based interval could perform even worse.}. An informal falsification test for whether such a monotonic transformation exists is whether the median of the simulated estimates is aligned with the point estimate; if it is not, there is no monotonic transformation that will yield a symmetric quantile interval with the desired coverage.

\section{Related software}\label{related-software}

Similar functionality exists in the \pkg{CLARIFY} package in Stata\footnote{Despite the similar name, the R package \CRANpkg{clarify} and the Stata package \pkg{CLARIFY} differ in several ways, one of which is that the estimates reported by \CRANpkg{clarify} in R are those computed using the original model coefficients, whereas those reported by \pkg{CLARIFY} in Stata are those computed as the average of the simulated distribution. The R implementation avoids the ``simulation-induced bias'' described by \citet{raineyCarefulConsiderationCLARIFY2023}.} \citep{tomzClarifySoftwareInterpreting2003} and used to be available in the \pkg{Zelig} R package \citep{imaiCommonFrameworkStatistical2008a}, though there are differences in these implementations. \CRANpkg{clarify} provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). \CRANpkg{clarify} relies on and can be seen as a companion to the \CRANpkg{marginaleffects} package \citep{marginaleffectsforthcoming}, which offers similar functionality but primarily uses the delta method for calculating uncertainty (though simulation-based inference is supported in a more limited capacity as well).

@@

For a running example, we will use the \texttt{lalonde} dataset in the \CRANpkg{MatchIt} package \citep{JSSv042i08}, which contains data on 614 participants enrolled in a job training program or sampled from a survey \citep{dehejiaCausalEffectsNonexperimental1999}. The treatment variable is \texttt{treat} and the outcome is \texttt{re78}, and all other variables are confounders. Although the original use of this dataset was to estimate the effect of \texttt{treat} on \texttt{re78}, we will use it more generally to demonstrate all of \CRANpkg{clarify}'s capabilities. In addition, we will use a transformation of the outcome variable to demonstrate applications to nonlinear models, for which the benefits of simulation-based inference are more apparent.

@@

\texttt{plot()} displays a density plot of the resulting estimates across the simulations, with markers identifying the point estimate (computed using the original model coefficients as recommended by \citet{raineyCarefulConsiderationCLARIFY2023}) and, optionally, uncertainty bounds (which function like confidence or credible interval bounds). The arguments to \texttt{plot()} are below:

@@

\begin{itemize}
\item
  \texttt{x} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be plotted if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be plotted.
\item
  \texttt{ci} -- whether to display lines at the uncertainty bounds. The default is \texttt{TRUE} to display them.
\item
  \texttt{level} -- if \texttt{ci} is \texttt{TRUE}, the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when \texttt{method} (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- if \texttt{ci} is \texttt{TRUE}, the method used to compute the bounds. Allowable methods include the Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). The Normal approximation involves multiplying the standard deviation of the estimates (i.e., which functions like the standard error of the sampling distribution) by the critical Z-statistic computed using \texttt{(1-level)/2} to create a symmetric margin of error around the point estimate. The default is \texttt{"quantile"} to instead use quantile-based bounds.
\item
  \texttt{reference} -- whether to display a normal density over the plot for each estimate and an indicator line for the median of the estimate. The default is \texttt{FALSE} to omit them.
\end{itemize}

@@

\begin{center}\includegraphics[alt={A plot of the density of the simulated predicted probability for participant PSID1, overlaid with a Normal distribution and a reference line for the median of the estimates}]{figures/plot1-1} \end{center}

@@

\begin{itemize}
\item
  \texttt{object} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be displayed if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be displayed.
\item
  \texttt{level} -- the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when \texttt{method} (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- the method used to compute the uncertainty bounds. Allowable methods include a Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). See \texttt{plot()} above.
\item
  \texttt{null} -- an optional argument specifying the desired null value in a hypothesis test for the estimates. If specified, a p-value will be computed using either a standard Z-test (if \texttt{method} is \texttt{"wald"}) or an inversion of the uncertainty interval (described below). The default is not to display any p-values.
\end{itemize}

@@

\begin{center}\includegraphics[alt={A plot of the density of the simulated difference between two regression coefficients, overlaid with a Normal distribution and a reference line for the median of the estimates.}]{figures/plot2-1} \end{center}

@@

\section{\texorpdfstring{Wrappers for \texttt{sim\_apply()}: \texttt{sim\_setx()}, \texttt{sim\_ame()}, and \texttt{sim\_adrf()}}{Wrappers for sim\_apply(): sim\_setx(), sim\_ame(), and sim\_adrf()}}\label{wrappers-for-sim_apply-sim_setx-sim_ame-and-sim_adrf}

@@

\texttt{sim\_setx()} provides an interface to compute predictions at representative and user-supplied values of the predictors. For example, we might want to know what the effect of treatment is for a ``typical'' individual, which corresponds to the contrast between two model-based predictions (i.e., one under treatment and one under control for a unit with ``typical'' covariate values). This functionality mirrors the \texttt{setx()} and \texttt{setx1()} functionality of \pkg{Zelig} (which is where its name originates) and provides similar functionality to functions in \CRANpkg{modelbased}, \CRANpkg{emmeans}, \CRANpkg{effects}, and \CRANpkg{ggeffects}.

@@

\begin{verbatim}
#>                         treat married      age     educ  race nodegree     re74
#> treat = 0, re75 = 0         0       0 27.36319 10.26873 black        1 4557.547
#> treat = 1, re75 = 0         1       0 27.36319 10.26873 black        1 4557.547
#> treat = 0, re75 = 20000     0       0 27.36319 10.26873 black        1 4557.547
#> treat = 1, re75 = 20000     1       0 27.36319 10.26873 black        1 4557.547
#>                          re75
#> treat = 0, re75 = 0         0
#> treat = 1, re75 = 0         0
#> treat = 0, re75 = 20000 20000
#> treat = 1, re75 = 20000 20000
\end{verbatim}

@@

\begin{center}\includegraphics[alt={Two plots of the densities of the simulated predictions of the outcome for control and treated units, one corresponding to re74 = 0 and the other to re75 = 20000.}]{figures/plot3-1} \end{center}

@@

\begin{center}\includegraphics[alt={A plot of the predicted values of the outcome across levels of re75 from 0 to 20000 with a line for control units and a line for treated units. Each line has an uncertainty band around it.}]{figures/plot4-1} \end{center}

@@

Using predicted values and effects at representative values is one way to summarize regression models, but another way is to compute average adjusted predictions (AAPs), contrasts of AAPs, and average marginal effects (AMEs). The definitions for these terms may vary and the names for these concepts differ across sources, but here we define AAPs as the average of the predicted values for all units after setting one predictor to a chosen value, and we define AMEs for binary predictors as the contrast of two AAPs and for continuous predictors as the average of instantaneous rate of change in the AAP corresponding to a small change in the predictor from its observed values across all units\footnote{In \CRANpkg{marginaleffects}, AAPs are computed using \texttt{avg\_predictions()}, AMEs for binary variables are computed using \texttt{avg\_comparisons()}, and AMEs for continuous variables are computed using \texttt{avg\_slopes()}. AAPs are sometimes known as average ``counterfactual'' predictions.} \citep{longRegressionModelsCategorical2014}.

@@

Here, we will use \texttt{sim\_ame()} to compute the AME of \texttt{treat} just among those who were treated (in causal inference, this is known as the average treatment effect in the treated, or ATT \citep{greiferChoosingCausalEstimand2023}). We will request our estimate to be on the risk ratio scale.

@@

\begin{center}\includegraphics[alt={A plot of the average dose-response function for the outcome against age. The line has an uncertainty band around it.}]{figures/plot8-1} \end{center}

@@

\begin{center}\includegraphics[alt={A plot of the average marginal effect function for the outcome against age. The line has an uncertainty band around it.}]{figures/plot9-1} \end{center}

@@

\section{Transforming and combining estimates}\label{transforming-and-combining-estimates}

@@

\section{\texorpdfstring{Using \CRANpkg{clarify} with multiply imputed data}{Using  with multiply imputed data}}\label{using-with-multiply-imputed-data}

Multiple imputation is a popular method of estimating quantities of interest in the presence of missing data and involves creating multiple versions of the original dataset each with the missing values imputed with estimates from an imputation model. Simulation-based inference in multiply imputed data is relatively straightforward. Simulated coefficients are drawn from the model estimated in each imputed dataset separately, and then the simulated coefficients are pooled into a single set of simulated coefficients. In Bayesian terms, this would be considered ``mixing draws'' and is the recommended approach for Bayesian analysis with multiply imputed data \citep{zhouNoteBayesianInference2010}.

@@

\section{Comparison to other packages}\label{comparison-to-other-packages}

>>>>

\abstract{%
Simulation-based inference is an alternative to the delta method for computing the uncertainty around regression post-estimation (i.e., derived) quantities such as average marginal effects, average adjusted predictions, and other functions of model parameters. It works by drawing model parameters from their joint distribution and estimating quantities of interest from each set of simulated values, which form a simulated ``posterior'' distribution of the quantity from which confidence intervals can be computed. \texttt{clarify} provides a simple, unified interface for performing simulation-based inference for any user-specified derived quantities as well as wrappers for common quantities of interest. \texttt{clarify} supports a large and growing number of models through its interface with the \texttt{marginaleffects} package and provides native support for multiply imputed data.
}

@@

The delta method involves computing a first-order Taylor series approximation to the variance of the derived quantity, and standard Wald-based inference relies on computing quantiles based on the Normal distribution and using them to compute p-values and confidence intervals. \CRANpkg{clarify} implements an alternative to the delta method---simulation-based inference---which involves simulating a ``posterior'' distribution of the derived quantities. Simulation-based inference does not require understanding Taylor series or the calculus that underlies it, which can make it more palatable to nontechnical audiences and easier to learn for students without necessarily sacrificing statistical performance (King, Tomz, and Wittenberg 2000; Zelner 2009). Some studies have found that simulation-based inference performs as well or better than the delta method for computing derived quantities (i.e., with respect to achieving close to nominal coverage for confidence intervals), especially for complicated derived quantities and in smaller samples (MacKinnon, Lockwood, and Williams 2004; Hole 2007; Herron 1999). Its empirical performance has been particularly well studied in the context of mediation analysis, in which the quantities of interest are products and ratios of regression coefficients, where it has been shown to perform well relative to the delta method due to the non-Normality of these quantities (Tofighi and MacKinnon 2016; Preacher and Selig 2012).

The methodology \CRANpkg{clarify} relies on was developed by Krinsky and Robb (1986) and is described in King, Tomz, and Wittenberg (2000) and Herron (1999). Simulation-based inference involves taking draws from a specified joint distribution of model parameters, computing derived quantities from these draws, and collecting the derived quantities in a ``posterior'' distribution, from which uncertainty measures (standard errors and confidence intervals) can be computed. This method assumes the model parameters are drawn from a multivariate Normal (or T) distribution with means at the estimated values and covariance equal to the asymptotic covariance matrix of the estimated values, a standard assumption motivated by the central limit theorem that underlies usual inference on the original model parameters. Arriving at the posterior distribution does not require taking any derivatives or making any approximations beyond those usually used for inference on model parameter estimates, except for the approximation due to Monte Carlo error induced by sampling from a finite number of simulations (which can always be reduced by increasing the number of draws at the cost of increased computing time).

The nonparametric bootstrap is another alternative to the delta method for inference that does not require its analytic approximations (Efron and Tibshirani 1986); bootstrapping typically involves re-sampling individuals from the sample, fitting the model in each bootstrap sample, and computing the quantity of interest from each model. Although bootstrapping tends to work well in practice, especially for complex and non-Normal estimators, refitting the model repeatedly can be prohibitively time-consuming and computationally expensive, especially for complicated models or large datasets. Simulation-based inference only requires the model to be fit once, and the simulations involve taking draws from a distribution produced from the single set of estimated parameters, making it much quicker in practice and allowing the user to capitalize on the already valid estimation of the model parameters. Methods for computing valid confidence intervals in cases when the quantity of interest has a complicated distribution are better developed when bootstrapping, however (Efron and Tibshirani 1986).

More formally, we fit a regression model \(y_i = f(x_i; \beta)\), such as a linear or other generalized linear model with model coefficients \(\beta\). We assume\[
\hat{\beta} \sim \text{MVN}(\beta, \Sigma_{\hat{\beta}})
\]where \(\hat{\beta}\) is the vector of estimates of \(\beta\) and \(\Sigma_{\hat{\beta}}\) is their asymptotic covariance matrix. We define a function \(\tau(\beta)\) that represents a quantity of interest derived from the model parameters, and compute its estimate \(\widehat{\tau(\beta)}\) as \(\tau(\hat{\beta})\). To perform simulation-based inference, we take \(M\) draws \(\tilde{\beta}^{(j)}\) for \(j\in(1, \dots, M)\) from a multivariate Normal distribution with mean vector \(\mu = \hat{\beta}\) and covariance \(\Sigma = \hat{\Sigma}_{\hat{\beta}}\), where \(\hat{\Sigma}_{\hat{\beta}}\) is an estimate of the asymptotic covariance matrix of the parameter estimates. We use the distribution of \(\tilde{\tau}=\tau(\tilde{\beta})\) as the ``posterior'' distribution of \(\widehat{\tau(\beta)}\), and compute its variance as\[
\hat\sigma^2_{\widehat{\tau(\beta)}} = \frac{1}{M-1}\sum^M_{j=1}{(\tilde{\tau}^{(j)} - \bar{\tilde{\tau}})^2}
\]and quantile \(100(1-\alpha)\%\) confidence interval limits as \(\left[\tilde{\tau}_{(\frac{\alpha}{2})}, \tilde{\tau}_{(1-\frac{\alpha}{2})}\right]\) where \(\tilde{\tau}_{(q)}\) is the \(q\)th value of \(\tilde{\tau}\) when arranged in ascending order (i.e., the \(q\)th quantile of the empirical cumulative distribution function). Simulation-based Wald-type confidence intervals can be computed as \(\left[\tau(\hat{\beta}) + \hat\sigma_{\widehat{\tau(\beta)}} Z_{\frac{\alpha}{2}}, \tau(\hat{\beta}) + \hat\sigma_{\widehat{\tau(\beta)}} Z_{1-\frac{\alpha}{2}} \right]\), where \(Z_q\) is the \(q\)th quantile of a standard Normal distribution. The delta method-based Wald-type confidence intervals use this formula but with the first-order Taylor approximation to the asymptotic variance: \(\hat\sigma^2_{\widehat{\tau(\beta)}}=\nabla\tau(\hat{\beta}) \Sigma_{\hat{\beta}} \nabla\tau'(\hat{\beta})\), where \(\nabla\tau(\hat{\beta})\) is the gradient of \(\tau(\beta)\) with respect to \(\beta\) evaluated at \(\hat{\beta}\).

To compute a p-value for a hypothesis test involving the quantity of interest, i.e., \(\text{H}_0: \tau (\beta) = \tau_0\) with a given null value \(\tau_0\), we can invert the confidence interval (Thulin 2021); that is, we find the largest value of \(\alpha\) such that \(\tau_0\) is within the confidence interval and use that \(\alpha\) as the p-value for the test. For Wald-based inference (either using the simulation-based variance or delta method-based variance), this is equivalent to performing a standard two-sided Z-test using the test statistic \(Z=\left(\tau(\hat\beta)-\tau_0\right) / \hat\sigma_{\widehat{\tau(\beta)}}\). One benefit of using the quantile p-values for inference is that equivalent tests of the same hypothesis will always yield identical p-values; for example, testing the equality of two derived quantities will yield the same p-value when comparing the difference between the quantities against a null hypothesis of 0 and the ratio of the quantities against a null hypothesis of 1, as each of these hypothesis is true if and only if the other is true.

One would expect simulation-based quantile inference, simulation-based Wald inference, and delta method-based Wald inference to align when the posterior is Normally distributed around the estimate, in which case any discrepancies would be due to Monte Carlo error in the simulated values (and therefore would shrink with increasing draws). However, for low values of \(\alpha\), it may require many draws for the simulation-based intervals to stabilize; delta method-based intervals are not subject to this error. There are a few cases in which the results might diverge: in some cases, the first-order Taylor series approximation to the variance may be poor, though in practice the approximation error is small and shrinks quickly with increasing sample size. When the posterior distribution is non-Normal but symmetric around the estimate, the quantile intervals may be more accurate (i.e., in the sense of achieving closer to nominal coverage) because they do not rely on quantiles from the Normal distribution (Tofighi and MacKinnon 2016).

Another potential advantage quantile intervals can have over Wald-intervals is that when \emph{some} monotonic transformation of the estimate has a symmetric distribution centered around the transformed estimate, the quantile intervals can achieve correct coverage without requiring knowledge of \emph{which} transformation is required (Efron and Tibshirani 1986); this is true of the quantile-based p-values as well. When the distribution is not centered around the estimate and no monotonic transformation will make it so, though, neither quantile-based nor Wald-based intervals would be expected to perform well, and quantile intervals could yield even worse coverage than Wald-based intervals, a phenomenon that occurs in the context of bootstrapping {[}Efron and Tibshirani (1986){]}\footnote{We thank an anonymous reviewer for pointing out a scenario in which this could occur: for a quantity of interest with a right-skewed sampling distribution, one would prefer an estimate to the right of the quantity's true value to have a confidence interval skewed to left to capture the bulk of the sampling distribution, but in practice a quantile confidence interval would also be skewed to the right. While a symmetric Wald-based interval may not have adequate coverage, the quantile-based interval could perform even worse.}. An informal falsification test for whether such a monotonic transformation exists is whether the median of the simulated estimates is aligned with the point estimate; if it is not, there is no monotonic transformation that will yield a symmetric quantile interval with the desired coverage.

\section{Related Software}\label{related-software}

Similar functionality exists in the \pkg{CLARIFY} package in Stata\footnote{Despite the similar name, the R package \CRANpkg{clarify} and the Stata package \pkg{CLARIFY} differ in several ways, one of which is that the estimates reported by \CRANpkg{clarify} in R are those computed using the original model coefficients, whereas those reported by \pkg{CLARIFY} in Stata are those computed as the average of the simulated distribution. The R implementation avoids the ``simulation-induced bias'' described by Rainey (2023).} (Tomz, Wittenberg, and King 2003) and used to be available in the \texttt{Zelig} R package (Imai, King, and Lau 2008), though there are differences in these implementations. \CRANpkg{clarify} provides additional flexibility by allowing the user to request any derived quantity, in addition to providing shortcuts for common quantities, including predictions at representative values, average marginal effects, and average dose-response functions (described below). \CRANpkg{clarify} relies on and can be seen as a companion to the \CRANpkg{marginaleffects} package (Arel-Bundock, Greifer, and Heiss Forthcoming), which offers similar functionality but primarily uses the delta method for calculating uncertainty (though simulation-based inference is supported in a more limited capacity as well).

@@

For a running example, we will use the \texttt{lalonde} dataset in the \CRANpkg{MatchIt} package (Ho et al. 2011), which contains data on 614 participants enrolled in a job training program or sampled from a survey (Dehejia and Wahba 1999). The treatment variable is \texttt{treat} and the outcome is \texttt{re78}, and all other variables are confounders. Although the original use of this dataset was to estimate the effect of \texttt{treat} on \texttt{re78}, we will use it more generally to demonstrate all of \CRANpkg{clarify}'s capabilities. In addition, we will use a transformation of the outcome variable to demonstrate applications to nonlinear models, for which the benefits of simulation-based inference are more apparent.

@@

\texttt{plot()} displays a density plot of the resulting estimates across the simulations, with markers identifying the point estimate (computed using the original model coefficients as recommended by Rainey (2023)) and, optionally, uncertainty bounds (which function like confidence or credible interval bounds). The arguments to \texttt{plot()} are below:

@@

\begin{itemize}
\item
  \texttt{x} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be plotted if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be plotted.
\item
  \texttt{ci} -- whether to display lines at the uncertainty bounds. The default is \texttt{TRUE} to display them.
\item
  \texttt{level} -- if \texttt{ci} is \texttt{TRUE}, the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when \texttt{method} (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- if \texttt{ci} is \texttt{TRUE}, the method used to compute the bounds. Allowable methods include thea Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). The Normal approximation involves multiplying the standard deviation of the estimates (i.e., which functions like the standard error of the sampling distribution) by the critical Z-statistic computed using \texttt{(1-level)/2} to create a symmetric margin of error around the point estimate. The default is \texttt{"quantile"} to instead use quantile-based bounds.
\item
  \texttt{reference} -- whether to display a normal density over the plot for each estimate and an indicator line for the median of the estimate. The default is \texttt{FALSE} to omit them.
\end{itemize}

@@

\begin{center}\includegraphics{figures/plot1-1} \end{center}

@@

\begin{itemize}
\item
  \texttt{object} -- the \texttt{clarify\_est} object (the output of a call to \texttt{sim\_apply()}).
\item
  \texttt{parm} -- the names or indices of the quantities to be displayed if more than one was estimated in \texttt{sim\_apply()}; if unspecified, all will be displayed.
\item
  \texttt{level} -- the desired two-sided confidence level. The default is .95 so that that the bounds are at the .025 and .975 quantiles when method (see below) is \texttt{"quantile"}.
\item
  \texttt{method} -- the method used to compute the uncertainty bounds. Allowable methods include a Normal approximation (\texttt{"wald"}) or using the quantiles of the resulting distribution (\texttt{"quantile"}). See \texttt{plot()} above.
\item
  \texttt{null} -- an optional argument specifying the desired null value in a hypothesis test for the estimates. If specified, a p-value will be computed using either a standard Z-test (if \texttt{method} is \texttt{"wald"}) or an inversion of the uncertainty interval (described below). The default is not to display any p-values.
\end{itemize}

@@

\begin{center}\includegraphics{figures/plot2-1} \end{center}

@@

\section{\texorpdfstring{\texttt{sim\_apply()} Wrappers: \texttt{sim\_setx()}, \texttt{sim\_ame()}, \texttt{sim\_adrf()}}{sim\_apply() Wrappers: sim\_setx(), sim\_ame(), sim\_adrf()}}\label{sim_apply-wrappers-sim_setx-sim_ame-sim_adrf}

@@

\texttt{sim\_setx()} provides an interface to compute predictions at representative and user-supplied values of the predictors. For example, we might want to know what the effect of treatment is for a ``typical'' individual, which corresponds to the contrast between two model-based predictions (i.e., one under treatment and one under control for a unit with ``typical'' covariate values). This functionality mirrors the \texttt{setx()} and \texttt{setx1()} functionality of \texttt{Zelig} (which is where its name originates) and provides similar functionality to functions in \CRANpkg{modelbased}, \CRANpkg{emmeans}, \CRANpkg{effects}, and \CRANpkg{ggeffects}.

@@

\begin{verbatim}
#>                         treat married      age     educ  race nodegree     re74  re75
#> treat = 0, re75 = 0         0       0 27.36319 10.26873 black        1 4557.547     0
#> treat = 1, re75 = 0         1       0 27.36319 10.26873 black        1 4557.547     0
#> treat = 0, re75 = 20000     0       0 27.36319 10.26873 black        1 4557.547 20000
#> treat = 1, re75 = 20000     1       0 27.36319 10.26873 black        1 4557.547 20000
\end{verbatim}

@@

\begin{center}\includegraphics{figures/plot3-1} \end{center}

@@

\begin{center}\includegraphics{figures/plot4-1} \end{center}

@@

Using predicted values and effects at representative values is one way to summarize regression models, but another way is to compute average adjusted predictions (AAPs), contrasts of AAPs, and average marginal effects (AMEs). The definitions for these terms may vary and the names for these concepts differ across sources, but here we define AAPs as the average of the predicted values for all units after setting one predictor to a chosen value, and we define AMEs for binary predictors as the contrast of two AAPs and for continuous predictors as the average of instantaneous rate of change in the AAP corresponding to a small change in the predictor from its observed values across all units\footnote{In \CRANpkg{marginaleffects}, AAPs are computed using \texttt{avg\_predictions()}, AMEs for binary variables are computed using \texttt{avg\_comparisons()}, and AMEs for continuous variables are computed using \texttt{avg\_slopes()}. AAPs are sometimes known as average ``counterfactual'' predictions.} (Long and Freese 2014).

@@

Here, we will use \texttt{sim\_ame()} to compute the AME of \texttt{treat} just among those who were treated (in causal inference, this is known as the average treatment effect in the treated, or ATT (Greifer and Stuart 2023)). We will request our estimate to be on the risk ratio scale.

@@

\begin{center}\includegraphics{figures/plot8-1} \end{center}

@@

\begin{center}\includegraphics{figures/plot9-1} \end{center}

@@

\section{Transforming and Combining Estimates}\label{transforming-and-combining-estimates}

@@

\section{\texorpdfstring{Using \pkg{clarify} with Multiply Imputed Data}{Using  with Multiply Imputed Data}}\label{using-with-multiply-imputed-data}

Multiple imputation is a popular method of estimating quantities of interest in the presence of missing data and involves creating multiple versions of the original dataset each with the missing values imputed with estimates from an imputation model. Simulation-based inference in multiply imputed data is relatively straightforward. Simulated coefficients are drawn from the model estimated in each imputed dataset separately, and then the simulated coefficients are pooled into a single set of simulated coefficients. In Bayesian terms, this would be considered ``mixing draws'' and is the recommended approach for Bayesian analysis with multiply imputed data (Zhou and Reiter 2010).

@@

\section{Comparison to Other Packages}\label{comparison-to-other-packages}
